### 2014-10-18 fundamentals/impl.tex
cleaning up implementation, adding content

A modular, agent-based is an ideal approach for solving complicated
physics-dependent supply chain problems involving material routing, facility
deployment, regional and institutional hierarchies.

@@

\Cyclus is primarily written in \texttt{C++}, and support is currently provided
for Linux-based (including Ubuntu and OSX) platforms. Furthermore, the core
infrastructure and related archetypes are free and open-source, BSD-3-clause
licsened. No part of \Cyclus is proprietary or based on commercial, off the
shelf (COTS) software. Accordingly, it can be easily deployed on large compute
systems, such as high-throughput computing (HTC) systems.

@@

A key innovation that has previously not been implemented in nuclear fuel cycle 
simulators in the literature is implementing this generic API and modular 
architecture into a suite of dynamically loadable ``plugin'' libraries.
This is implemented largely through a clean API and a modern build system.

@@

This approach allows efficient, targeted contribution to the ecosystem of libraries.  The 
scientist-developer can focus on generating a model within their
sphere of expertise, while relying on the model contributions of others to fill 
in the other technologies.  This strategy also allows individual developers to
explore different levels of complexity within their archetypes, including
wrapping other simulation tools as loadable libraries within the \Cyclus
framework.

@@

This last benefit of dynamically-loadable libraries addresses
another goal of \Cyclus: ubiquity amongst its potential user base. By
engineering \Cyclus to easily handle varying levels of complexity, a single
simulation engine can be used by both users interested in big-picture policy
questions as well as users focused on more detailed, technical
analyses.

@@

A key compontent of the \Cyclus kernel architecture is the ability for
implementations of actors in a given simulation to be easily
interchanged. Critically, this novel functionality enables the comparison
between agent implementations. For example, a low-physics-fidelity
implementation of a reactor can be compared to an implementation with higher
physics fidelity, allowing an analyst to discern the effect of reactor physical
fidelity on a given fuel cycle.

@@

\Cyclus provides a novel representation of entities in the nuclear fuel cycle,
including facilities, institutions managing those facilities, and regions. While
some simulators (\TODO{cite DESAE}) have provided a notion of static regional
effects, \Cyclus allows for both regions and institutions to be first-class
actors in simulated fuel cycles.

The basic interaction models for each entity is implemented by a corresponding
archetype class in \Cyclus, i.e., the \Class{Region} class, \Class{Institution}
class, and \Class{Facility} class. Archetype developers can then build on the
provided functionality by subclassing the appropriate class.

\Cyclus implements a Region-Institution-Facility (RIF) relationship through the
parent-child hierarchy described in \S \ref{sec:interchangeability}, where
regions are the parents of institutions which are, in turn, the parents of
facilities. In other words, RIF hierarchies form a directed acyclic graph (DAG),
with regions as root nodes and facilities as leaf nodes.

@@

All operations performed on every resource object (splitting, combining,
decay, etc.) are tracked in detail.  This information includes the creating
agent for when resources are newly created and introduced into the simulation.
The parentage of each resource is also tracked. This makes it possible to
follow the travel history of discrete resource objects as they are transferred
between agents.

The \Cyclus kernel has built-in experimental support for decay calculations.
Materials "remember" the time since their last decay and agents are free to
invoke the decay function on them as desired to decay them to the current
simulation time.  Cyclus can currently operate in 2 decay modes with 2 other
modes likely to be added in future releases:

@@

In large simulations there may be many material objects changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in Cyclus have some special features.  Compositions in Cyclus are
immutable.  This immutability allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that don't change the composition, any new resulting
materials hold a reference to the same, originating composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new an originating compositions. The next time the originating
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

>>>>

A modular, agent-based is an ideal approach for solving complex, coupled,
physics-dependent supply chain problems involving material routing, facility
deployment, regional and institutional hierarchies. Additionally, the choice to 
build \Cyclus on open source libraries in modern programming languages enables 
both remote and mutiprocess execution on a number of platforms.

Many fuel cycle simulators rely on \gls{COTS} and Windows-only software that limits 
performance on resource computing infrastructures. This limitation severely 
increases the wall-clock time necessary to conduct parameterized sensitivity 
analyses and other multi-simulation studies.

\Cyclus, on the other hand, is primarily written in \texttt{C++} and relies on 
libraries supported by Linux and UNIX (including Ubuntu and OSX) platforms. 
Furthermore, the core infrastructure and related archetypes are free and 
open-source, BSD-3-clause licsened. No part of \Cyclus is proprietary or based 
on \gls{COTS}, Windows-based software. Accordingly, it can be easily deployed 
on large compute systems, such as high-throughput computing (HTC) systems.

@@

@@

A key innovation in \Cyclus is implementing this generic API and modular 
architecture into a suite of dynamically loadable ``plugin'' libraries.  Though 
common in modern software architecture, such a ``plugin'' pattern has 
previously not been implemented in nuclear fuel cycle simulators in the 
literature.  In \Cyclus, this is made possible primarily by the clean, 
object-oriented \gls{API} and modern build system already described.

@@

This approach encourages efficient, targeted contribution to the ecosystem of 
archetype libraries.  The 
scientist-developer can focus on generating an archetype model within their
sphere of expertise, while relying on the contributions of others to fill 
in the other technologies in the simulation.  This strategy also allows individual developers to
explore different levels of complexity within their archetypes, including
wrapping other simulation tools as loadable libraries within the \Cyclus
framework.

@@

This last benefit of dynamically-loadable libraries addresses
another goal of \Cyclus: ubiquity amongst its potential user base. By
engineering \Cyclus to easily handle varying levels of complexity, a single
simulation engine can be used by both users interested in big-picture policy
questions as well as users focused on more detailed, technical
analyses. They merely choose their preferred level of fidelity from amongst 
available libraries.

@@

Due to this ``plugin'' architecture
implementations of actors in a given simulation can be easily
interchanged. Critically, this novel functionality enables the comparison
between agent implementations. For example, a low-physics-fidelity
implementation of a reactor can be compared to an implementation with higher
physics fidelity, allowing an analyst to discern the effect of reactor physical
fidelity on a given fuel cycle.

@@

\Cyclus provides a novel representation of entities in the nuclear fuel cycle 
which mimics the reality in international nuclear power production
including facilities, institutions managing those facilities, and regions. While
some simulators (\TODO{cite DESAE}) have provided a notion of static regional
effects, \Cyclus allows for both regions and institutions to be first-class
actors in simulated fuel cycles.

The fundamental interactions for each entity are implemented in a corresponding
archetype class in \Cyclus, i.e., the \Class{Region} class, \Class{Institution}
class, and \Class{Facility} class. Archetype developers can then build on the
provided functionality by inheriting from the appropriate class.

\Cyclus implements a \gls{RIF} relationship through the
parent-child hierarchy described in \S \ref{sec:interchangeability}, where
regions are the parents of institutions which are, in turn, the parents of
facilities. In other words, RIF hierarchies form a directed acyclic graph (DAG),
with regions as root nodes and facilities as leaf nodes.

@@

All operations performed on every resource object (splitting, combining,
decay, etc.) are tracked in detail.  This information includes the creating
agent for when resources are newly created and introduced into the simulation.
The parentage of each resource is also tracked. This makes it possible to
follow the travel history of discrete resource objects as they are transferred
between agents and powerfully impacts the ability to model 
non-proliferation-related simulation goals.

The \Cyclus kernel has built-in experimental support for decay calculations.
Materials ``remember'' the time since their last decay and agents are free to
invoke the decay function on them as desired to decay them to the current
simulation time.  Cyclus can currently operate in 2 decay modes with 2 other
modes likely to be added in future releases:

@@

In large simulations there may be many material objects changing frequently, on 
the day or month timescale.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in Cyclus have some special features.  Compositions in Cyclus are
immutable.  This immutability allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that don't change the composition, any new resulting
materials hold a reference to the same, originating composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new an originating compositions. The next time the originating
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs. \TODO{do we want a diagram helping show this composition caching?}

### 2014-10-18 fundamentals/intro.tex
cleaning up implementation, adding content

\Cyclus utilizes an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it takes advantage of modern computing resources.  
Furthermore, agent-based modeling provides an analyst a more natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context for example, an analyst can design an agent with reactor-related logic 
and a separate agent with fuel fabrication-related logic. 
 Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bonafide actors in the simulation.  The two archetype libraries, therefore, need 
not have any shared knowledge of one another for them to be building blocks in 
the same simulation.

>>>>

\Cyclus utilizes an \gls{ABM} paradigm.  \gls{ABM} is modern, sophisticated, 
true to reality, and it takes advantage of modern computing resources.  
Furthermore, agent-based modeling provides an analyst a more natural 
perspective from which to model a given problem. In the nuclear fuel cycle 
context for example, an analyst can design an agent with reactor-related logic 
entirely independently from a separate agent with fuel fabrication-related 
logic. Defining the behavior of both agents according to the 
\gls{API} contract is sufficient for them to interact with one another as 
bonafide actors in the simulation.  The two archetype libraries, therefore, need 
not have any shared knowledge of one another for them to be building blocks in 
the same simulation.
